### 2022-03-15 abstract.tex
finish paper 1
Finish paper 2 revisions

The online world presents new avenues for the spread of information in addition to the traditional word of mouth effects of social networks. It also provides new big data sources for social science research because they are bigger, cheaper, and already available. In my dissertation, I combine these two areas of inquiry by examining the differences between the social network contagion of ideas (push) and information seeking (pull) and their relation to the creation of social norms. My first article tests the construct validity of the Google Search Trends source of big data as an indicator of three different cases, namely cultural attitudes, disease prevalence and voting behavior, using pairwise correlations . My second article establishes how individuals perform information search, whether through interpersonal networks or online, using my original survey data of 900 Adults in the United States. My final article builds on these by investigating how new health-behavior social norms form in the face of assortative signals through associative cognitive diffusion. This research will be a contribution to methodology of computational social science as it tests the reliability of a big data source, questions what behavioral processes lead to macro-level outcomes recorded in big data sources, and uses novel combinations of data sources and methodology to investigate questions of sociological interest. This research will also contribute to the sociology study of information technologies in the digital age, medical sociology, and social networks.

>>>>

% at least 150 words, generally no more than 500 words

This dissertation focuses on how people search for information and how people
rely on this information to inform their health-behaviors and develop social
norms. Information can be shared and promoted by others
through information campaigns and consumed by an individual, called information
push, or information can be intentionally sought out by individuals, 
called information pull, and their relation to the creation of social norms.

The first article tests the construct validity of the Google Search Trends source,
an online indicator of computer-mediated information search. I attempt to
validate Google Search Trends for use as indicators of attitudes, disease prevalence
and political preferences using eight different validated data sources. %TODO double check number
I fail to find correlation among any of the Google Trends tested and their
validated indicators and show that there is no external validity of
Google Trends for these uses and social scientists will find no replacement for
high quality survey data with Google Trends. Instead, we must only use 
Google Trends to demonstrate interest or attention.

Knowing that Google trends data only encompasses a small portion of the
information-seeking done by modern humans, my second article 
is motivated by the research question: How do computer-mediated or interpersonal 
information-seeking strategies vary across populations? Using original survey data
of 948 Americans, I investigate their experiences seeking out information about 
the Covid-19 vaccines among the following information seeking vehicles: 
personal connection, doctor, social networking site, online forum, and online search engine. 
I find little evidence that online search is more utilized than
seeking social support from personal network connections or health professionals
as I hypothesize based on uses and gratifications theory. I find that different
exposure points and information search vehicles hold real world consequences
through their associations with Covid-19 vaccination rates and intentions, as
information from a doctor increases the Covid-19 vaccination uptake while
receiving information from a Social Networking Site like Facebook or Twitter was
associated with lower odds of vaccination.

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty using the cases of
both stay-at-home rates and vaccination rates as responses to public health recommendations to mitigate the Covid-19 pandemic. Using theories of associative diffusion and the integrated
theoretical framework of norms, I test models of behavioral adaption to public health
recommendations as well as patterns of complex social contagion using linear
mixed effects models. These models show that complex contagion is a valid 
framework for the social contagion of new norms during Covid-19. Importantly, I find
a novel moderating effect of signal discordance, the contextual diversity of
signals received by an ego. This paper shows that the contagion process is not
fully understood without looking at the context of each exposure point within the
range of exposures one experiences.

The introductory chapter provides a summary of the research and an explanation of how this
research contributes to the sociology, social science, and society.

### 2022-03-15 ch1.tex
Finish paper 2 revisions

The importance of reputable information and misinformation was clear
in the recent and ongoing Covid-19 pandemic. While the vaccines against the
Covid-19 virus prevent severe illness, hospitalization, and death, only 60\% of
Arizonans over 12 years old are fully vaccinated against the virus. We know that
people exposed to misinformation are much less likely to receive vaccinations
against the virus \citep{loombaMeasuringImpactCOVID192021}. It is not an
exaggeration to claim that misinformation has been deadly for Arizona
\citep{pathakInfodemicsCOVID19Role2020, greene_murphy21}.

@@

Max Weber, another founding architect of sociology saw social norms differently.
Weber distinguished social behavior from social action, which was an action a
person takes through subjective understanding and interpretation of actions of
others \citeyearpar{weber1978economy}. Weber therefore saw social norms through
the lens of the subjective meanings behind every action. Actions that were not
considered by an individual were not the focus of Weber. New spontaneous norms
that are emerging in a society under upheaval, like the health behaviors that
emerged during the Covid-19 pandemic, emerged with new subjective meanings and
cognitive associations that held consequences for the individuals who adopted
the social actions.

@@

\hyperlink{paper-2}{\subsection{Article 2: Information-Seeking Behaviors for Information on Covid-19 Vaccinations}}

Knowing that Google trends data only encompasses a small portion of the
information-seeking done by modern humans, it is important to
investigate what leads individuals to search for a question online using
a search engine versus more
traditional means of information gathering, such as network activation
(asking someone in their social network for informational support). 
Previous research has greatly failed to distinguish between the activation of
information seeking behaviors online and offline. Using theories of social
support and uses and gratifications theory, I investigate the factors associated
with each information seeking vehicle: personal connection,
doctor, social networking site, online forum, and online search engine. I use
original survey data of 948 Americans to investigate their experiences seeking out
information about the Covid-19 vaccines.

This paper aims to address two main questions:
\begin{itemize}
  \item How do computer-mediated or interpersonal information-seeking strategies vary
across populations?
  \item How does the information search vehicle utilized affect Covid-19 Vaccination uptake? 
\end{itemize}

In this article, I find little evidence that online search is more utilized than
seeking social support from personal network connections or health professionals
as I hypothesize based on uses and gratifications theory. I find evidence that
the categorizations of vehicles queried in this survey are indeed conceptually
different from each other and that the utilization of different vehicles varies
by demography and information exposure points. Finally, I find that different
exposure points and information search vehicles hold real world consequences
through their associations with Covid-19 vaccination rates and intentions, as
information from a doctor increases the Covid-19 vaccination uptake while
receiving information from a Social Networking Site like Facebook or Twitter was
associated with lower odds of vaccination.

\hyperlink{paper-3}{\subsection{Article 3: Social Norms under Uncertain Times: A dynamic study of Stay-At-Home and Vaccination Rates During the Covid-19 Pandemic}}

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty using the cases of
both stay-at-home rates and vaccination rates as responses to public health recommendations to mitigate the Covid-19 pandemic. Using theories of associative diffusion and the integrated
theoretical framework of norms, I test models of behavioral adaption to public health
recommendations as well as patterns of complex social contagion using linear
mixed effects models.

These models show that complex contagion is a valid 
framework for the social contagion of new norms during Covid-19. Importantly, I find
a novel moderating effect of signal discordance, the contextual diversity of
signals received by an ego. This paper shows that the contagion process is not
fully understood without looking at the context of each exposure point within the
range of exposures one experiences.

@@

% search as an agentic process
This research will make contributions to multiple subfields and literatures. 
First and foremost, these articles contribute to the sociological literature.
The spread of information has long been investigated in the discipline (diffusion, mass
communication, social influence). However, it is almost always researched from a structural
perspective which disregards individual agency and choice in the
acquisition of new information. This dissertation adopts the second perspective,
centering people as active agents in their search for information aligned
with their own preferences and gratifications instead of just
passive receivers of information signals.
Taking this agentic perspective is critical in the study of information
diffusion when combined with cognitive perspectives like that of
Goldberg and Stein (2018). How people search for information determines
the information that they find. The information they uncover is filtered
through cognitive biases and predispositions to how they interpret and
act upon the newfound information. Searching for informational support
through social network ties theoretically will uncover potentially
different information than that would be uncovered through online web
search and will change how people are organized and how people behave.
\hyperlink{paper-2}{Article 2} importantly shows that 
the vehicles used to search for information have a significant impact on the choice to get vaccinated against Covid-19, a real-world consequence to the minutiae of individual information search.

>>>>

The importance of reputable information and misinformation was clear
in the recent and ongoing COVID-19 pandemic. While the vaccines against the
COVID-19 virus prevent severe illness, hospitalization, and death, only 60\% of
Arizonans over 12 years old are fully vaccinated against the virus. We know that
people exposed to misinformation are much less likely to receive vaccinations
against the virus \citep{loombaMeasuringImpactCOVID192021}. It is not an
exaggeration to claim that misinformation has been deadly for Arizona
\citep{pathakInfodemicsCOVID19Role2020, greene_murphy21}.

@@

Max Weber, another founding architect of sociology saw social norms differently.
Weber distinguished social behavior from social action, which was an action a
person takes through subjective understanding and interpretation of actions of
others \citeyearpar{weber1978economy}. Weber therefore saw social norms through
the lens of the subjective meanings behind every action. Actions that were not
considered by an individual were not the focus of Weber. New spontaneous norms
that are emerging in a society under upheaval, like the health behaviors that
emerged during the COVID-19 pandemic, emerged with new subjective meanings and
cognitive associations that held consequences for the individuals who adopted
the social actions.

@@

\hyperlink{paper-2}{\subsection{Article 2: Information-Seeking Behaviors for Information on COVID-19 Vaccinations}}

Knowing that Google trends data only encompasses a small portion of the
information-seeking done by modern humans, it is important to
investigate what leads individuals to search for a question online using
a search engine versus more
traditional means of information gathering, such as network activation
(asking someone in their social network for informational support). 
Previous research has greatly failed to distinguish between the activation of
information seeking behaviors online and offline. Using theories of social
support and uses and gratifications theory, I investigate the factors associated
with each information seeking vehicle: personal connection,
doctor, social networking site, online forum, and online search engine. I use
original survey data of 948 Americans to investigate their experiences seeking out
information about the COVID-19 vaccines.

This paper aims to address two main questions:
\begin{itemize}
  \item How do computer-mediated or interpersonal information-seeking strategies vary
across populations?
  \item How does the information search vehicle utilized affect COVID-19 Vaccination uptake? 
\end{itemize}

In this article, I find little evidence that online search is more utilized than
seeking social support from personal network connections or health professionals
as I hypothesize based on uses and gratifications theory. I find evidence that
the categorizations of vehicles queried in this survey are indeed conceptually
different from each other and that the utilization of different vehicles varies
by demography and information exposure points. Finally, I find that different
exposure points and information search vehicles hold real world consequences
through their associations with COVID-19 vaccination rates and intentions, as
information from a doctor increases the COVID-19 vaccination uptake while
receiving information from a Social Networking Site like Facebook or Twitter was
associated with lower odds of vaccination.

\hyperlink{paper-3}{\subsection{Article 3: Social Norms under Uncertain Times: A dynamic study of Stay-At-Home and Vaccination Rates During the COVID-19 Pandemic}}

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty using the cases of
both stay-at-home rates and vaccination rates as responses to public health recommendations to mitigate the COVID-19 pandemic. Using theories of associative diffusion and the integrated
theoretical framework of norms, I test models of behavioral adaption to public health
recommendations as well as patterns of complex social contagion using linear
mixed effects models.

These models show that complex contagion is a valid 
framework for the social contagion of new norms during COVID-19. Importantly, I find
a novel moderating effect of signal discordance, the contextual diversity of
signals received by an ego. This paper shows that the contagion process is not
fully understood without looking at the context of each exposure point within the
range of exposures one experiences.

@@

% search as an agentic process
This research will make contributions to multiple subfields and literatures. 
First and foremost, these articles contribute to the sociological literature.
The spread of information has long been investigated in the discipline (diffusion, mass
communication, social influence). However, it is almost always researched from a structural
perspective which disregards individual agency and choice in the
acquisition of new information. This dissertation adopts the second perspective,
centering people as active agents in their search for information aligned
with their own preferences and gratifications instead of just
passive receivers of information signals.
Taking this agentic perspective is critical in the study of information
diffusion when combined with cognitive perspectives like that of
Goldberg and Stein (2018). How people search for information determines
the information that they find. The information they uncover is filtered
through cognitive biases and predispositions to how they interpret and
act upon the newfound information. Searching for informational support
through social network ties theoretically will uncover potentially
different information than that would be uncovered through online web
search and will change how people are organized and how people behave.
\hyperlink{paper-2}{Article 2} importantly shows that 
the vehicles used to search for information have a significant impact on the choice to get vaccinated against COVID-19, a real-world consequence to the minutiae of individual information search.

### 2022-03-15 paper1.tex
3 article summaries
Finish paper 2 revisions

@@

Building on prior research outlining the various issues with different sources of big data
\citep{boydCriticalQuestionsBig2012,lazerIssuesConstructValidity2015},
this paper evaluates the construct validity of Google Search Trends as an indicator of three different
cases, namely attitudes, disease prevalence and political preferences.
These three cases will be tested using attitudinal indicators from the %NORC General Social Survey, 
CDC \citeyearpar{vaches_data} and New York Times \citeyearpar{mask_data}, 
health indicators from the CDC \citeyearpar{suic_data}, 
US cases rates of Covid-19 from The New York Times \citeyearpar{covid_data},  
and political indicators from the historical US Presidential Election results \citeyearpar{pres_data}, 
and the American National Election Survey \citeyearpar{anes_data}. This paper will contribute to the
creation of methodological norms and standards of how to use Google
Trends as a big data source for societal research and serve as a
critical inquiry into the adoption of big data without a critical eye
for the ecological validity of the sources.

@@

Google Search Trends have been utilized to estimate disease prevalence
and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
the Covid-19 pandemic \citep{jimenez_etal20, jimenezCOVID19SymptomGoogle2020, limEstimatingInformationSeekingBehaviour2020, mavraganiCOVID19PredictabilityUnited2020, nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021, mingUnderstandingHealthCommunication2021},
other research has investigated Google Trends as indicators of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021, carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, 
vaccination uptake \citep{dalumhansenEnsembleLearnedVaccination2016}, 
obesity \citep{sarigulNowcastingObesityUsing2014}, 
and even insomnia \citep{zittingGoogleTrendsReveal2020},
to cover a few examples. For a partial review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to \citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment \citep{dechoudhurySeekingSharingHealth2014}.

@@

While Google Trends has not been widely used or validated by social scientists,
I am not the first author to investigate its validity.
In contrast to the papers mentioned above, there are researchers who have cautioned against
the wide usage of Google Search Trends. For instance, \citet{asseoTrackingCOVID19Using2020} 
concludes in their research on Covid-19 case rates and Google searches that the correlation 
is often spurious and due to confounding. They even explain that "searches would
result not only from self-symptoms, but also from interest elicited by
media coverage" \citep[][p.1]{asseoTrackingCOVID19Using2020}.

\citet{rovetta21} also investigates the reliability of Covid-19 searches to infer 
the quality of the data; the author finds that there are various issues with missing data in 
smaller regions and significant variation in correlations between the sampled regions throughout
the study period. Rovetta indicates that this is a cause for concern and is undoubtedly due to more 
serious causes than behavioral changes among the population or changes in the underlying product algorithm.

@@

I also use U.S. Covid-19 rates to validate health and disease related topics. I
retrieve U.S. county-level Covid-19 rates from \citet{covid_data}, who compile this
data based on reports from state and local health agencies. It is widely
acknowledged that there are biases in this data due to inconsistencies and
availability in testing as well as different community propensity to test
\citep{gu22, cdc20a} However, it is the best measure we have of actual case rates.
Case Rate is measured as number of cases per 100,000 population. Observations
vary from a minimum of zero to a maximum of 1460.46 for each Monday from January 27, 2020, through
December 27, 2021. There are 285,986 cases across 3,136 counties and 101
dates. Missing data were interpolated using \texttt{na.approx()} \citep{zoo}.

@@

In addition to attitudinal measures, I attempt to validate Google Trends for
uses in the measurements of health indicators. The first indicator I assess is
Covid-19 case rates from 2020 through 2021. The Pearson correlation results
indicate negligible to weak relationships between the search terms and the
actual case rates across time and place. The repeated measures correlation
coefficient, a more reliable and less biased measure for longitudinal
correlations, reveals similarly weak results with a maximum correlation of 0.32
between 'smell loss' and rates of Covid-19. To further investigate the relationship, I
include the three trends in random intercept hierarchical linear models. Model 1
of table \ref{tab:covid_analysis} indicates that the trends themselves are able
to explain about 20\% of the variation in Covid-19 rates ($r^2$ = 0.201). Few
of the demographic variables influence Covid-19 rates in models 2 or 3,
though there is some suppression for median income, unemployment rates,
and population density. Model 3 has an $r^2$ of 0.200. I conclude that the
Google Trends are slightly related to Covid-19 rates, but that this relationship
is weak, and we should not attempt to use Google Trends as an indicator of actual
health outcomes.

>>>>

% TODO: should I add in some figures to outline these relationships?

@@

Building on prior research outlining the various issues with different sources of big data
\citep{boydCriticalQuestionsBig2012,lazerIssuesConstructValidity2015},
this paper evaluates the construct validity of Google Search Trends as an indicator of three different
cases, namely attitudes, disease prevalence and political preferences.
These three cases will be tested using attitudinal indicators from the %NORC General Social Survey, 
CDC \citeyearpar{vaches_data} and New York Times \citeyearpar{mask_data}, 
health indicators from the CDC \citeyearpar{suic_data}, 
US cases rates of COVID-19 from The New York Times \citeyearpar{covid_data},  
and political indicators from the historical US Presidential Election results \citeyearpar{pres_data}, 
and the American National Election Survey \citeyearpar{anes_data}. This paper will contribute to the
creation of methodological norms and standards of how to use Google
Trends as a big data source for societal research and serve as a
critical inquiry into the adoption of big data without a critical eye
for the ecological validity of the sources.

@@

Google Search Trends have been utilized to estimate disease prevalence
and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
the COVID-19 pandemic \citep{jimenez_etal20, jimenezCOVID19SymptomGoogle2020, limEstimatingInformationSeekingBehaviour2020, mavraganiCOVID19PredictabilityUnited2020, nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021, mingUnderstandingHealthCommunication2021},
other research has investigated Google Trends as indicators of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021, carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, 
vaccination uptake \citep{dalumhansenEnsembleLearnedVaccination2016}, 
obesity \citep{sarigulNowcastingObesityUsing2014}, 
and even insomnia \citep{zittingGoogleTrendsReveal2020},
to cover a few examples. For a partial review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to \citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment \citep{dechoudhurySeekingSharingHealth2014}.

@@

While Google Trends has not been widely used or validated by social scientists,
I am not the first author to investigate its validity.
In contrast to the papers mentioned above, there are researchers who have cautioned against
the wide usage of Google Search Trends. For instance, \citet{asseoTrackingCOVID19Using2020} 
concludes in their research on COVID-19 case rates and Google searches that the correlation 
is often spurious and due to confounding. They even explain that "searches would
result not only from self-symptoms, but also from interest elicited by
media coverage" \citep[][p.1]{asseoTrackingCOVID19Using2020}.

\citet{rovetta21} also investigates the reliability of COVID-19 searches to infer 
the quality of the data; the author finds that there are various issues with missing data in 
smaller regions and significant variation in correlations between the sampled regions throughout
the study period. Rovetta indicates that this is a cause for concern and is undoubtedly due to more 
serious causes than behavioral changes among the population or changes in the underlying product algorithm.

@@

I also use U.S. COVID-19 rates to validate health and disease related topics. I
retrieve U.S. county-level COVID-19 rates from \citet{covid_data}, who compile this
data based on reports from state and local health agencies. It is widely
acknowledged that there are biases in this data due to inconsistencies and
availability in testing as well as different community propensity to test
\citep{gu22, cdc20a} However, it is the best measure we have of actual case rates.
Case Rate is measured as number of cases per 100,000 population. Observations
vary from a minimum of zero to a maximum of 1460.46 for each Monday from January 27, 2020, through
December 27, 2021. There are 285,986 cases across 3,136 counties and 101
dates. Missing data were interpolated using \texttt{na.approx()} \citep{zoo}.

@@

In addition to attitudinal measures, I attempt to validate Google Trends for
uses in the measurements of health indicators. The first indicator I assess is
COVID-19 case rates from 2020 through 2021. The Pearson correlation results
indicate negligible to weak relationships between the search terms and the
actual case rates across time and place. The repeated measures correlation
coefficient, a more reliable and less biased measure for longitudinal
correlations, reveals similarly weak results with a maximum correlation of 0.32
between 'smell loss' and rates of COVID-19. To further investigate the relationship, I
include the three trends in random intercept hierarchical linear models. Model 1
of table \ref{tab:covid_analysis} indicates that the trends themselves are able
to explain about 20\% of the variation in COVID-19 rates ($r^2$ = 0.201). Few
of the demographic variables influence COVID-19 rates in models 2 or 3,
though there is some suppression for median income, unemployment rates,
and population density. Model 3 has an $r^2$ of 0.200. I conclude that the
Google Trends are slightly related to COVID-19 rates, but that this relationship
is weak, and we should not attempt to use Google Trends as an indicator of actual
health outcomes.
